---
title: "Lisa-results"
format: html
editor: visual
---

## Depression trend

Interesting results of our analyses in the existing dataset included the trend analysis of the share of depression diagnoses in the population from 2014:

```{r}
#| echo: false

library(readr)
library(tidyverse)
library(ggrepel)
library(ggplot2)
library(dplyr)

# Load data
rki_data <- read_tsv("GBE_Indikatoren_nichtuebertragbarer_Erkrankungen.tsv")

# Region_ID is read as character due to some international rows mixed in;
# convert to numeric and drop those non-German rows
rki_data <- rki_data %>%
  mutate(Region_ID = as.numeric(Region_ID)) %>%
  filter(!is.na(Region_ID))

# Filter to the relevant subset:
# depression indicator, age-adjusted, across education levels,
# both genders combined, all age groups combined
rki_data_lisa_vis <- rki_data %>%
  filter(Indikator_ID == 2040202) %>% # depression rows
  filter(Bildung_Casmin_ID == 0) %>% # across education levels
  filter(Geschlecht_ID == 0) %>% # both genders combined
  mutate(Jahr = as.numeric(Zeitraum_Name))

# Germany-wide (i.e. not broken down by region)
rki_data_lisa_vis_1 <- rki_data_lisa_vis %>%
  filter(Alter_ID == "00+") %>% # all age groups
  filter(Region_ID == 0) %>%
  filter(Standardisierung_ID == 3) # age-adjusted data

# Plot: Germany-wide trend with confidence interval ribbon
p <- ggplot(rki_data_lisa_vis_1, aes(x = Jahr, y = Wert)) +
  geom_ribbon(aes(ymin = Unteres_Konfidenzintervall, ymax = Oberes_Konfidenzintervall), fill = "#0072B2", alpha = 0.15) +
  geom_line(color = "#0072B2", linewidth = 1) +
  geom_point(color = "#0072B2", size = 2) +
  geom_text_repel(
    aes(label = Jahr),
    size = 3.2,
    color = "grey30",
    nudge_y = 0.3,
    segment.size = 0.2,
    min.segment.length = 0
  ) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(labels = function(x) paste0(x, " %")) +
  labs(
    title = "Depression diagnoses in Germany have risen over time",
    subtitle = "Share of population with diagnosed depression, by year -
    width of graph represents 95% confidence interval",
    x = "Year",
    y = NULL,
    caption = "Data source: Informationssystem der Gesundheitsberichterstattung (GBE-Bund)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey30"),
    plot.caption = element_text(color = "grey50", hjust = 0)
  )

p
```

The code is directly taken from rki_Skript.R. Relevant parts have been extracted and commented by Claude.

For the sake of having a more complicated plot to practice plot interactivity on, I created the following plot:

```{r}
library(plotly)

rki_data_lisa_vis_2 <- rki_data_lisa_vis %>%
  filter(Standardisierung_Name == "beobachtet") %>% # observed (not age-adjusted, since age is our grouping var)
  filter(Region_ID == 0) %>% # Germany-wide, not regional
  filter(Alter_ID != "00+") # exclude the "all ages" aggregate

# Plot: depression trend over time, one line per age group
p_age <- ggplot(rki_data_lisa_vis_2, aes(x = Jahr, y = Wert, color = Alter_Name, group = Alter_Name)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(labels = function(x) paste0(x, " %")) +
  labs(
    title = "Depression diagnoses in Germany over time, by age group",
    subtitle = "Share of population with diagnosed depression, by year and age group",
    x = "Year",
    y = NULL,
    color = "Age group",
    caption = "Data source: Informationssystem der Gesundheitsberichterstattung (GBE-Bund)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey30"),
    plot.caption = element_text(color = "grey50", hjust = 0)
  )

ggplotly(p_age)
```
