---
title: "alt_proposal.qmd"
format: html
editor: visual
---

Source and License:

This data is sourced from the the health report of the Robert-Koch-Institut (<https://github.com/robert-koch-institut/Gesundheitsberichterstattung_-_Daten_zu_nichtuebertragbaren_Erkrankungen?tab=CC-BY-4.0-1-ov-file#readme>) and it is publicly available under the CC-BY-4.0 license.

What is in it?

The data set contains health data on non-infectious diseases in the german population. It has 29 columns and 53,036 rows. Each row constitutes an observation of the monthly prevalence of a specific health behavior or illness (physical or mental) in a sample. The variable "Indikator_ID" gives a identifying number to each indicator (ergo the illness or behavior). The variable "Indikator_Name" is the name of the indicator. A detailled description can be found in the ReadMe under indicator on the website (see link above). The variable "Wert" contains the prevalence of the indicator in percent. The simple size is in the variable "Stichprobe", the region is under "Region_Name" and

Research questions

1.  question proposal: How did depressive symptoms develop over time and do depressive symptoms correlate with social support (Indikator_ID = `1010301`) in different regions of Germany?
2.  question proposal: What are differences in depressive symptoms between people of different educational levels (Bildung_Casmin_Name)? And did depressive symptoms of the most educated group decrease?

Context and target audience:

**Initial data analysis:**

Glimpse output:

```         
Rows: 53,036
Columns: 29
$ Indikator_ID                  <dbl> 1010301, 1010301, 1010301, 1010301, 1010301, 1010301, 1010301, 10103…
$ Indikator_Name                <chr> "Soziale Unterstützung (ab 18 Jahre)", "Soziale Unterstützung (ab 18…
$ Kennzahl_ID                   <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,…
$ Kennzahl_Name                 <chr> "Soziale Unterstützung", "Soziale Unterstützung", "Soziale Unterstüt…
$ Kennzahl_Definition           <chr> "Anteil der Erwachsenen mit starker sozialer Unterstützung in %", "A…
$ Zeitraum_ISO                  <chr> "2009-01-01--2009-12-31", "2009-01-01--2009-12-31", "2009-01-01--200…
$ Zeitraum_Name                 <chr> "2009", "2009", "2009", "2009", "2009", "2009", "2009", "2009", "200…
$ Geschlecht_ID                 <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
$ Geschlecht_Name               <chr> "Gesamt", "Gesamt", "Gesamt", "Gesamt", "Gesamt", "Gesamt", "Gesamt"…
$ Alter_ID                      <chr> "00+", "00+", "00+", "00+", "00+", "00+", "00+", "00+", "00+", "00+"…
$ Alter_Name                    <chr> "Alle Altersgruppen", "Alle Altersgruppen", "Alle Altersgruppen", "A…
$ Region_ID                     <chr> "00", "00", "00", "00", "00", "00", "00", "00", "01", "01", "02", "0…
$ Region_Name                   <chr> "Deutschland", "Deutschland", "Deutschland", "Deutschland", "Deutsch…
$ Bildung_Casmin_ID             <dbl> 0, 0, 1, 1, 2, 2, 3, 3, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
$ Bildung_Casmin_Name           <chr> "Gesamt", "Gesamt", "niedrig", "niedrig", "mittel", "mittel", "hoch"…
$ GISD_ID                       <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ GISD_Name                     <chr> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ Berufliche_Qualifikation_ID   <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ Berufliche_Qualifikation_Name <chr> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ Standardisierung_ID           <dbl> 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0, 3, 0,…
$ Standardisierung_Name         <chr> "beobachtet", "altersstandardisiert", "beobachtet", "altersstandardi…
$ Wert                          <dbl> 32.6, 32.4, 28.4, 30.0, 34.8, 32.9, 35.8, 35.8, 34.4, 34.4, 34.0, 33…
$ Unteres_Konfidenzintervall    <dbl> 31.8, 31.6, 26.9, 28.4, 33.6, 31.7, 34.4, 34.1, 30.1, 30.1, 28.7, 28…
$ Oberes_Konfidenzintervall     <dbl> 33.4, 33.2, 30.0, 31.7, 35.9, 34.2, 37.3, 37.5, 38.9, 39.0, 39.8, 38…
$ Fälle                         <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ Stichprobe                    <dbl> 20418, 20418, 4963, 4963, 10319, 10319, 5104, 5104, 737, 737, 459, 4…
$ Unsicherheit                  <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
$ Anmerkung                     <chr> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ Datenstand                    <date> 2024-11-20, 2024-11-20, 2024-11-20, 2024-11-20, 2024-11-20, 2024-11…
```

Summary of key variables

Since the data includes a lot of different indicators which are not seperated into columns but spread over different rows, one can only really calculate the amount of data points at this point (basically, all unique entries of Indikator_Name could be an own column). Calculating means or standard deviations across the indicators does not provide real meaning here. The data is not transformed to make unique entries of Indikator_Name into columns as we are only choosing a small subset of values of Indikator_Name and matching rows would be hard (for example because of differing age categories across indicators). It would also inflate the number of columns.

Missing values:

```{r}
library(readr)

df <- read_tsv("GBE_Indikatoren_nichtuebertragbarer_Erkrankungen.tsv")

key_vars <- c(
 "Kennzahl_Name", "Zeitraum_Name", "Region_Name",
 "Geschlecht_Name", "Alter_Name", "Bildung_Casmin_Name",
 "Wert", "Unteres_Konfidenzintervall", "Oberes_Konfidenzintervall",
 "Stichprobe", "Unsicherheit"
)

df %>%
 select(all_of(key_vars)) %>%
 summarise(across(everything(), ~ sum(is.na(.)))) %>%
 pivot_longer(everything(), names_to = "Variable", values_to = "N_Missing") %>%
 mutate(Pct_Missing = round(N_Missing / nrow(df) * 100, 1)) %>%
 print(n = Inf)
```

There are a lot of values missing for confidence intervals. This is partly explained by "Unsicherheit" being to low or the indicator not being an estimate based on a sample. That also explains the missing values in sample.

Gender:

```{r}
df %>%
   group_by(Geschlecht_Name) %>%
   summarise(
     N      = n(),
   ) %>%
   print()
```

There seems to be roughly similar depth in data for women and men.

Age:

```{r}
df %>%
   group_by(Alter_Name) %>%
   summarise(
     N      = n(),
   ) %>%
   print()
```

Since categories are inconsistent, this is difficult to interpret.

Education:

```{r}
df %>%
   group_by(Bildung_Casmin_Name) %>%
   summarise(
     N      = n(),
   ) %>%
   print()
```

Region_Name:

```{r}
df %>%
   group_by(Region_Name) %>%
   summarise(
     N      = n(),
   ) %>%
   print(n=60)
```

This shows that the data contains German states (majority), European countires and other (West, Ost, ...). For our analyses we expect to mainly use German data on a federal and state level.

Stichprobe:

```{r}
df %>%
  summarise(
    Mean   = round(mean(Stichprobe, na.rm = TRUE), 2),
    SD     = round(sd(Stichprobe, na.rm = TRUE), 2),
    Median = round(median(Stichprobe, na.rm = TRUE), 2),
    Min    = round(min(Stichprobe, na.rm = TRUE), 2),
    Max    = round(max(Stichprobe, na.rm = TRUE), 2)  
  ) %>%
  print()

```

Unsicherheit:

```{r}
df %>%
   group_by(Unsicherheit) %>%
   summarise(
     N      = n(),
   ) %>%
   print()
```

Unsicherheit is fairly small for most rows but should be checked for relevant indicators regardless.

Quality issues:

-one variable has the wrong data type (Zeitraum_Name) RESOLVED: By creating Start_Beobachtungszeitraum and Ende_Beobachtungszeitraum from Zeitraum_ISO we got the information we need and it is often times more precise than in Zeitraum_Name. Converting Zeitraum_Name to date would result in some form of data loss as it is in varying formats (e.g. 2024/25, 2009).

```{r}
unique(df$Zeitraum_Name)
```

\- many variables are double although integer would be sufficient ADDRESSED in rki_Skript

-there is no variable uniquely identifing each row/observation ADDRESSED in rki_Skript

-some variables have more missing values then existing values (e.g. Fälle, Unteres_Konfidenzintervall)

-because of the use of indicators system the variable "Indikator_ID" basically contains different attributes/variables; that is unhandy to work with

-Age categories are inconsistent, so depending on what indicators we want to combine into an analysis, we might not be able to use Age as a control variable

-Bildung_Casmin_Name is not available across all indicators so we might not be able to use that as a control variable depending on the analysis.

-Sample size is really small for some indicators. We will have to keep in mind to check the sample size for relevant indicators and calculate power.

Solution:

Analysis plan:

1\.

Variables:

- Indikator_ID

- time variable depends on whether the time intervals are entire years, let's see:

```{r}
rki_data_1_1 <- df %>% 
  filter(Indikator_ID == '2040202' | Indikator_ID == '1010301')

unique(rki_data_1_1$Zeitraum_ISO)
```

Let's see whether Zeitraum_Name is sufficient then:

```{r}
unique(rki_data_1_1$Zeitraum_Name)
```

-\> it is, so it is easiest to go with that column

- Geschlecht_ID and Alter_ID as controlling variables

- Region_ID

- Standardisierung_ID: There are age-adjusted and raw data. As we, in a part of our analyses, want to control for age, we should only use the age-adjusted data if the age ranges are rather wide. So let's look at that:

```{r}
unique(rki_data_1_1$Alter_Name)
```

-\> they are, so we will focus on age-adjusted data (Standardisierung_ID == 3)

- Stichprobe

- Unteres_Konfidenzintervall

- Oberes_Konfidenzintervall

- Wert

- Unsicherheit

Analytical approach:

1.  Data cleaning
2.  Data transformation
3.  See if all data can be used (Unsicherheit)
4.  See if there is a similar Stichprobengröße for the different subgroups
5.  Correlation between age-adjusted depression and social support over time
6.  Zoom in on potential differences for gender (so seperate correlations for women and men), same for regions
7.  Calculate linear regression with depression as dependent variable and social support, age, gender, and region as independent variables – Edit: Unfortunately, the data only allowed for an analysis with social support with region and another analysis with social support, age and gender
8.  Calculate the analyses with Inverse-Variance Weighting

Cleaning / transformation steps needed: General data cleaning (data types), transform data so that one year is one row (match indicators)

Uncertainties:

- Some sub-groups might get pretty small - if the subgroups vary a lot in size, the Inverse-Variance Weighting might be skewed. But we will cross that bridge when we come to it.

- To keep the analytical complexity in a range that I can handle, I will not look at education more closely. It is possible, however, that education moderates the relationship between depression and social support.

2\.

Variables:

Analytical approach:

Cleaning / transformation steps needed:

Uncertainties:

Workflow organization:

Code style: tidyverse, enforced with styler linter (not included in the project yet, WOP)

Packages: number of packages should be kept at a minimum to avoid conflicts between packages but every collaborator can use new packages at their own may.

Planned packages: Git workflow: people should work on individual branches, named after the author, to reduce conflicts wherever it is practical. Exceptions can be made after getting permission from the branch creator or when meeting up in one place; continue the common practice to let the partner confirm the pull request

Gitignore: Artefacts from running code will be added to a gitignore file, though it is unclear at this stage what exactly we will include as we do not know the entire requirements of our project just yet
